---
title: "simpleRegress"
author: "Racheal Adeneye"
date: "April 22, 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
mHEADLEN <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(mHEADLEN)
summary.lm(mHEADLEN)
```

According to the observations, the residual standard error is 67.81, and our adjusted R-squared is 0.69. 

At the moment, the best model is WEIGHT~NECK(RSE: 43.91) followed by WEIGHT~HEADLEN(67.81). 


```{r}
mHEADWTH <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(mHEADWTH)
summary.lm(mHEADWTH)
```

In his plot, the residual standard error is 76.41 and the adjusted R squared is 0.6064. 

Best to worst by RSE    
NECK RSE: 43.91     
HEADLEN RSE: 67.81       
HEADWTH RSE: 76.41 


```{r}
mLENGTH <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(mLENGTH)
summary.lm(mLENGTH)
```

In this plot, we can see that the RSE is 61.83 and the adjusted R-squared is 0.7423.

Best to worst by RSE      
NECK RSE: 43.91      
LENGTH RSE: 61.83     
HEADLEN RSE: 67.81         
HEADWTH RSE: 76.41

```{r}
mCHEST <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(mCHEST)
summary.lm(mCHEST)
```

In this final plot, we see that the residual standard error is 33.08, and the adjusted R-squared is 0.962.   


Best to worst by RSE     
CHEST RSE: 33.08    
NECK RSE: 43.91      
LENGTH RSE: 61.83     
HEADLEN RSE: 67.81         
HEADWTH RSE: 76.41  
